---
title: "Country Profile"
output:
  html_document:
    toc: true
params:
  my_iso3: "KEN"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# import required packages
library(tidyverse)
library(countrycode)
library(janitor)
library(scales)

# set default theme for plots
theme_set(theme_minimal())
```

# Data

The file `hiv_rates.csv` contains a tidy dataset with five columns:

1. The ISO3 code of the country that reported data.
1. The year for which data was reported.
1. The percentage of HIV-positive individuals age 15-49.
1. The percentage of HIV-positive females age 15-24.
1. The percentage of HIV-positive males age 15-24.

```{r import-data}
# import data and clean up
hiv <- read_csv("data/hiv_rates.csv") %>%
  clean_names() %>%
  mutate(country_name = countrycode(iso3, "iso3c", "country.name"))
```

```{r customize, include = FALSE}
# filter data for individual county specified in parameters
my_iso3 <- params$my_iso3

# get the single row
single_iso <- hiv %>%
  filter(iso3 == my_iso3)

# identify the country by name
my_country <- single_iso %>%
  distinct(country_name) %>%
  pull(country_name)

# get the most recent year
most_recent <- single_iso %>%
  summarize(max_year = max(year)) %>%
  pull()
```

# Country Report: `r my_country`

For which years do we have data for `r my_country`?

```{r years}
single_iso %>%
  distinct(year)
```

How has HIV incidence changed over time in `r my_country`?

```{r hiv-over-time}
# use stringr::str_glue() to paste in the specific country name
plot_title <- str_glue("Country: {my_country}")

single_iso %>%
  ggplot(aes(x = year, y = hiv_pct_15_to_49)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(labels = label_percent()) +
  labs(
    title = plot_title,
    subtitle = "HIV prevalence, age 15-49",
    x = "Year",
    y = NULL
  )

single_iso %>%
  ggplot(aes(x = year)) +
  geom_line(aes(
    y = hiv_pct_15_to_24_female,
    color = "Female"
  )) +
  geom_line(aes(
    y = hiv_pct_15_to_24_male,
    color = "Male"
  )) +
  scale_y_continuous(labels = label_percent()) +
  labs(
    title = plot_title,
    subtitle = "HIV prevalence, age 15-24",
    x = "Year",
    y = NULL,
    color = NULL
  )
```

How does `r my_country` compare in `r most_recent`?

```{r hiv-now, fig.height = 14}
plot_title <- glue::glue("Year: {most_recent}")

hiv %>%
  filter(year == most_recent) %>%
  ggplot(aes(x = fct_reorder(country_name, hiv_pct_15_to_49), y = hiv_pct_15_to_49)) +
  geom_col() +
  geom_col(data = single_iso %>% filter(year == most_recent), fill = "dodgerblue") +
  labs() +
  scale_y_continuous(labels = label_percent()) +
  coord_flip() +
  labs(
    title = plot_title,
    x = NULL,
    y = "HIV prevalence, age 15-49"
  )
```
